---
title: "The course workflow"
author: "Jeffrey Boichuk"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{The course workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# Nothing in this vignette calls the X API when it is built. Every chunk is
# shown, none is run.
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This article is the routine you follow once a week for the brand you are
tracking. Set up once, pull what is new, save it, unfold it into tables,
join them, and check the bill. Each step is short, and the order matters:
save before you unfold, and read the cost line before you spend.

## Set up once

Put your bearer token in `.Renviron` (`usethis::edit_r_environ()` opens
it) and restart R:

```
X_BEARER_TOKEN=AAAA...
```

Then look the brand up once. A handle costs one user read ($0.010) every
time a reader turns it into an id, so save the id to a small file and pass
`user_id` from then on. Keep the id as text: as a number it loses digits.

```{r setup-brand}
library(xapir)
library(dplyr)

brand <- get_users_by_usernames("Tesla")
#> Reading up to 1 users, about $0.01.

readr::write_csv(select(brand, user_id, username), "data-raw/brand.csv")
```

Every later week starts by reading that file back. `readr` keeps the id as
text on its own; `read.csv()` would turn it into a number.

```{r read-brand}
brand <- readr::read_csv("data-raw/brand.csv", col_types = "cc")
brand_id <- brand$user_id
```

## The weekly pull

Ask for only what is new. `since_id` takes the id of the newest post from
last week's pull, so a post is never paid for twice on purpose, and
`max_posts` caps the worst case.

```{r pull}
# NULL in week one, so the pull starts from the newest post
last_id <- if (file.exists("data-raw/last-post-id.rds")) {
  readRDS("data-raw/last-post-id.rds")
}

pages <- get_timeline(
  user_id   = brand_id,
  since_id  = last_id,
  max_posts = 500
)
#> Reading up to 500 posts, about $2.50. Set max_posts to change this.
#> Finished getting posts on page 1
#> Read 38 posts, about $0.19.
```

The first line is the most the call can cost; the last is what it did
cost, because a week rarely fills the cap. Save the raw pages to a dated
file before touching anything, and record the newest id for next week.

```{r save}
saveRDS(pages, sprintf("data-raw/tesla-%s.rds", Sys.Date()))

newest <- extract_post(pages, include_referenced_posts = FALSE)
saveRDS(newest$post_id[which.max(newest$created_at)], "data-raw/last-post-id.rds")
```

Reading a file back is free; reading the timeline again is not.

## Unfold

Each `extract_*()` function turns the saved pages into one table. None
calls the API, so run them as often as you like.

```{r unfold}
pages <- readRDS("data-raw/tesla-2026-09-25.rds")

post         <- extract_post(pages)
user         <- extract_user(pages)
post_media   <- extract_post_media(pages)
post_hashtag <- extract_post_hashtag(pages)
post_url     <- extract_post_url(pages)
```

Week over week, stack the new rows on the old ones. A post can appear in
two pulls, since the tables include the posts a timeline quoted, replied
to or reposted, and a post quoted this week may have been read last week.
`distinct()` keeps one copy.

```{r stack}
all_post <- bind_rows(readRDS("data/post.rds"), post) |>
  distinct(post_id, .keep_all = TRUE)
saveRDS(all_post, "data/post.rds")

all_user <- bind_rows(readRDS("data/user.rds"), user) |>
  distinct(user_id, .keep_all = TRUE)
saveRDS(all_user, "data/user.rds")

all_hashtag <- bind_rows(readRDS("data/post_hashtag.rds"), post_hashtag) |>
  distinct(post_id, hashtag, .keep_all = TRUE)
saveRDS(all_hashtag, "data/post_hashtag.rds")
```

The same two lines work for `post_media` (on `post_id` and `media_id`)
and `post_url` (on `post_id` and `url`). In week one there is no old file
yet, so `bind_rows()` gets only the new table.

## Join

Every post table carries `post_id`, and `user` carries `user_id`, the
author. Which hashtags did the brand use, and who wrote each post?

```{r join}
all_post |>
  inner_join(all_hashtag, by = "post_id") |>
  left_join(all_user, by = "user_id") |>
  select(created_at, username, hashtag, like_count) |>
  arrange(desc(like_count))
```

A first engagement summary: likes per post by type. Reposts carry `NA` for
likes, replies, quotes and bookmarks by design, because those numbers
belong to the original post, so they drop out of the mean rather than
pulling it towards zero.

```{r engagement}
all_post |>
  filter(user_id == brand_id) |>
  group_by(post_type) |>
  summarise(
    posts      = n(),
    mean_likes = mean(like_count, na.rm = TRUE),
    .groups    = "drop"
  )
```

## The conversation around the brand

`get_mentions()` returns the posts that name the account. It takes
`since_id` too, so it appends the same way: pull, save the dated file,
unfold, stack, keep the newest id.

```{r mentions}
mention_pages <- get_mentions(
  user_id   = brand_id,
  since_id  = readRDS("data-raw/last-mention-id.rds"),
  max_posts = 500
)
saveRDS(mention_pages, sprintf("data-raw/tesla-mentions-%s.rds", Sys.Date()))

mention <- extract_post(mention_pages)
```

Before paying for a search, size it. A count is billed once per request
($0.005 for the last seven days), however many posts it covers.

```{r count}
get_recent_post_count("@Tesla -is:retweet", granularity = "day")
#> This request costs about $0.005.
```

If the count says 4,000 posts, a `max_posts = 500` pull reads the newest
500 for $2.50 and stops; narrow the query or raise the cap on purpose.

## Keeping the bill in view

End the week with two calls that cost nothing. `get_spend()` prices each
day's post reads at the post price and prints one line.

```{r spend}
get_spend(days = 7)
#> Read 1,430 posts in the last 7 days, about $7.15 at the post price.
```

That figure is a ceiling: it counts post reads only, and reads of your own
data or of the same post twice in a day bill for less. The balance itself
comes from `get_usage_credits()`.

```{r credits}
get_usage_credits()
```

Reading a saved file, unfolding it and joining the tables never touch the
API, so the bill moves only when you pull.
