---
title: "Getting Started"
author: "Jeffrey Boichuk"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# Nothing in this vignette calls the X API when it is built. Every chunk is
# shown, none is run.
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

xapir reads from and writes to X through the X API v2. You pull posts with
one function, save what comes back as a file, and unfold that file into
tidy tables with a second set of functions that never touch the API. X
bills every post it returns, so each reader tells you the most a call can
cost before it makes a request.

This guide covers four things: signing in, what a call costs, reading a
timeline and unfolding it into tables, and writing to X from your own
account.

## Install

The package is not on CRAN. Install a numbered release from GitHub:

```{r install}
# install.packages("pak")
pak::pak("Ivey-Business-School/xapir@v0.2.0")
library(xapir)
```

## Signing in

X issues two kinds of credentials, and the package uses both. Both come
from an app on the [X developer
portal](https://developer.x.com/en/portal/dashboard).

### A bearer token, for reading

The bearer token is a long string on your app's "Keys and tokens" page. It
lets you read public data: timelines, searches, profiles, lists and
trends. Put it in your `.Renviron` file, which `usethis::edit_r_environ()`
opens for you:

```
X_BEARER_TOKEN=AAAA...
```

Restart R. Every reader picks the token up from
`Sys.getenv("X_BEARER_TOKEN")`; you never pass it by hand. If the variable
is empty, the reader stops before it spends anything and says so.

### A client id, for acting as yourself

Anything that acts as your account needs a user token instead: posting,
deleting, reposting, bookmarking, following, muting and hiding a reply,
plus the readers that only make sense for the signed-in account
(`get_account_timeline()`, `get_bookmark()`, `get_liked_posts()`,
`get_liking_users()`, `get_repost_of_me()`, `get_my_user()`,
`get_blocking()` and `get_muting()`).

To get a user token the package needs your app's OAuth 2.0 client id, from
the same page in the portal. Add it beside the bearer token and restart R:

```
X_CLIENT_ID=your-client-id
```

In the portal, the app must have user authentication turned on, with
read-and-write permissions and `http://localhost:1410` as a callback URL.

The first time you call one of these functions a browser window opens and
asks you to approve the app. Approve it, come back to R, and the call
carries on. The token is saved to disk under `httr2::oauth_cache_path()`
in a folder named `xapir`, so you sign in once, not every session. When
the token expires it is refreshed in the background; if the refresh fails,
the browser opens again.

To sign in as a different account, delete that folder and call a user
function again:

```{r sign-out}
unlink(file.path(httr2::oauth_cache_path(), "xapir"), recursive = TRUE)
```

## What a call costs

X bills per item returned, not per request: US$0.005 per post and
US$0.010 per user at the prices of September 2026. Counts are free, and so
are fields, which is why every reader asks for every field by default.

Every reader prints the most it can spend before its first request, using
the `max_posts` (or `max_users`) cap:

```
#> Reading up to 500 posts, about $2.50. Set max_posts to change this.
```

A reader that pages through results also prints what it actually read at
the end, because a pull often stops before the cap:

```
#> Finished getting posts on page 1
#> Finished getting posts on page 2
#> Read 143 posts, about $0.72.
```

Two rules keep a mistake small. The cap must be a finite number, so
`max_posts = Inf` stops before the first request. And the last page is
trimmed, so a pull never holds more posts than you asked for.

Two details are worth knowing:

* A `username` costs one user read ($0.010) to turn the handle into an id.
  `get_timeline()`, `get_owned_list()` and `get_followed_lists()` take
  `user_id` instead, which skips that read. Keep ids as text,
  `"2244994945"`, because as numbers they lose digits.
* If X changes its prices, set
  `options(xapir.price_per_post = 0.006, xapir.price_per_user = 0.02)` in
  your `.Rprofile` and every message follows. The defaults are `0.005` and
  `0.01`.

## Read, save, unfold

### Read

`get_timeline()` reads an account's posts, newest first. Cap the pull, and
give a window with `start_time` and `end_time` when you want one.
`iso_8601()` turns a date into the format the API wants.

```{r read}
timeline <- get_timeline(
  username   = "Tesla",
  max_posts  = 500,
  start_time = iso_8601("2026-01-01"),
  end_time   = iso_8601("2026-02-01")
)
```

The result is a list of pages, each holding `data`, `includes` and `meta`
as the API returned them. It is not a table yet, on purpose: it holds
everything X sent, including the posts that were quoted or reposted and
the users, media, polls and places they refer to, so you never pay for the
same posts twice.

### Save

Save the raw pull straight away, before you touch it.

```{r save}
saveRDS(timeline, "tesla-2026-01.rds")
```

Every function in the rest of this guide reads from that file. Reading it
back is free; reading the timeline again is not.

```{r load}
timeline <- readRDS("tesla-2026-01.rds")
```

### Unfold

`extract_post()` turns the pages into one row per post with the same 22
columns every time.

```{r extract-post}
post <- extract_post(timeline)
post
```

The columns, in order: `created_at` (UTC; pass `tz = Sys.timezone()` for
local time), `text` (the whole post, even past 280 characters, with
`is_long_post` marking the long ones), `lang`, `possibly_sensitive`,
`article_title`, the six counts (`impression_count`, `like_count`,
`repost_count`, `quote_count`, `reply_count`, `bookmark_count`),
`reply_settings`, the ids of the post this one reposted, quoted or replied
to (`reposted`, `quoted`, `replied_to`), `in_reply_to_user_id`, `user_id`,
`conversation_id`, `post_type` (Thread, Post, Quote post, Reply or
Repost), `post_url` and `post_id`.

On a repost, `like_count`, `reply_count`, `quote_count`, `bookmark_count`
and `repost_count` are `NA`, because those numbers belong to the original
post. Its `impression_count` is its own.

Media come out with `extract_post_media()`: one row per media item per
post, so a post with three photos gives three rows. For a video or an
animated GIF, `url` is the mp4 with the highest bit rate the API offers.

```{r extract-media}
post_media <- extract_post_media(timeline)
post_media
```

The two tables join on `post_id`. Which videos drew the most views?

```{r join}
library(dplyr)

post |>
  select(post_id, created_at, text, like_count) |>
  inner_join(post_media, by = "post_id") |>
  filter(type == "video") |>
  arrange(desc(view_count))
```

## The twelve tables

Each `extract_*()` function reads the saved pull and returns one table.
Together they are the course's data model. None of them calls the API.

| Function                            | One row per                          | Key columns                                         |
|-------------------------------------|--------------------------------------|-----------------------------------------------------|
| `extract_post()`                    | post                                 | `post_id`, `user_id`, `text`, the counts            |
| `extract_user()`                    | user                                 | `user_id`, `username`, `followers_count`            |
| `extract_post_media()`              | media item per post                  | `post_id`, `media_id`, `type`, `url`, `view_count`  |
| `extract_post_url()`                | link per post                        | `post_id`, `expanded_url`, `title`, `image_url`     |
| `extract_post_mention()`            | mentioned account per post           | `post_id`, `username`, `user_id`                    |
| `extract_post_hashtag()`            | hashtag per post                     | `post_id`, `hashtag`                                |
| `extract_post_cashtag()`            | cashtag per post                     | `post_id`, `tag`                                    |
| `extract_post_context()`            | context annotation per post          | `post_id`, `domain_name`, `entity_name`             |
| `extract_post_entity_annotation()`  | named entity per post                | `post_id`, `normalized_text`, `type`, `probability` |
| `extract_post_poll_option()`        | poll option per post                 | `post_id`, `poll_id`, `label`, `votes`              |
| `extract_post_place()`              | tagged post                          | `post_id`, `place_id`, `full_name`, `country`       |
| `extract_post_edited_post_id()`     | earlier version of an edited post    | `post_id`, `edited_post_id`                         |

Eleven of the tables carry `post_id`, so any of them joins to the post
table on that column. `extract_user()` carries `user_id` instead and joins
to the post table's `user_id`, which is the author.

```{r join-user}
post |>
  left_join(extract_user(timeline), by = "user_id") |>
  select(username, created_at, text, like_count)
```

Two properties hold for every table. First, the columns are always the
same, even when there are no rows: a timeline with no polls gives a
zero-row poll table with the usual columns, so a script written on one
pull runs on the next. Second, each table includes the posts a timeline
quotes, replies to or reposts by default, so a quoted post's photo is in
the media table too. Pass `include_referenced_posts = FALSE` to keep only
the posts the endpoint returned. `extract_user()` has no such switch; it
lists every user in the pull.

The other post readers return the same list of pages and feed the same
tables: `get_recent_post()` (a search over the last seven days),
`get_post()` (up to 100 posts by id), `get_quote_post()`, `get_repost()`,
`get_mentions()`, `get_bookmark()`, `get_liked_posts()`,
`get_account_timeline()` and `get_repost_of_me()`. The user and list
readers, such as `get_users_by_usernames()` and `get_list_member()`,
return a tibble directly with the same 18 user columns as
`extract_user()`.

## Writing to X

Every write acts as the account that signed in, so each of these opens
the browser the first time. Each returns the API's `data` invisibly and
stops with X's own message when the request is refused. A rate limit or a
server error is retried, waiting as long as X asks.

```{r write}
# Post, and keep the id so you can delete it later
new_post <- create_post(text = "Hello from R!")
new_post$id

# Delete one or more of your own posts; one row per id
delete_post(post_ids = new_post$id)

# Repost and un-repost by post id
create_repost(post_id = "1234567890123456789")
delete_repost(post_id = "1234567890123456789")

# Bookmarks are always your own
create_bookmark(post_id = "1234567890123456789")
delete_bookmark(post_id = "1234567890123456789")

# Follow, unfollow, mute and unmute; the source is the account that signed in
follow_user(source_username = "your_handle", target_username = "XDevelopers")
unfollow_user(source_username = "your_handle", target_username = "XDevelopers")
mute_user(source_username = "your_handle", target_username = "XDevelopers")
unmute_user(source_username = "your_handle", target_username = "XDevelopers")

# Hide a reply to one of your posts, and show it again
hide_reply(reply_id = "1234567890123456789")
hide_reply(reply_id = "1234567890123456789", hidden = FALSE)
```

`delete_post()` deletes in batches of five with a fifteen-minute pause
between them, which stays inside X's deletion cap on every tier. Raise
`batch_size` on a paid tier to go faster.

## Where next

The [reference index](https://Ivey-Business-School.github.io/xapir/reference/index.html)
lists every function with its arguments and columns, grouped the same way
as this guide. The [release notes](https://Ivey-Business-School.github.io/xapir/news/index.html)
say what changed in each version and which column or argument names
moved.
