---
title: "Getting Started"
author: "Jeffrey Boichuk"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# Nothing in this vignette calls the X API when it is built. Every chunk is
# shown, none is run.
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

xapir reads from and writes to X through the X API v2. You pull posts with
one function, save what comes back as a file, and unfold that file into
tidy tables with a second set of functions that never touch the API. X
bills every post it returns, so each reader tells you the most a call can
cost before it makes a request.

This guide covers five things: signing in, what a call costs, reading a
timeline and unfolding it into tables, reading followers and lists, and
writing to X from your own account.

## Install

The package is not on CRAN. Install a numbered release from GitHub:

```{r install}
# install.packages("pak")
pak::pak("Ivey-Business-School/xapir@v0.2.0")
library(xapir)
```

## Signing in

X issues two kinds of credentials, and the package uses both. Both come
from an app on the [X developer
portal](https://developer.x.com/en/portal/dashboard).

### A bearer token, for reading

The bearer token is a long string on your app's "Keys and tokens" page. It
lets you read public data: timelines, searches, profiles, lists and
trends. Put it in your `.Renviron` file, which `usethis::edit_r_environ()`
opens for you:

```
X_BEARER_TOKEN=AAAA...
```

Restart R. Every reader picks the token up from
`Sys.getenv("X_BEARER_TOKEN")`; you never pass it by hand. If the variable
is empty, the reader stops before it spends anything and says so.

### A client id, for acting as yourself

Anything that acts as your account needs a user token instead: posting,
uploading media, deleting, liking, reposting, bookmarking, following,
muting, blocking, hiding a reply and managing lists, plus the readers
that only make sense for the signed-in account (`get_account_timeline()`,
`get_bookmark()`, `get_liked_posts()`, `get_liking_users()`,
`get_repost_of_me()`, `get_my_user()`, `get_blocking()`, `get_muting()`,
`get_pinned_lists()`, `get_post_analytics()`,
`get_personalized_trends()`, `search_users()` and
`search_communities()`).

To get a user token the package needs your app's OAuth 2.0 client id, from
the same page in the portal. Add it beside the bearer token and restart R:

```
X_CLIENT_ID=your-client-id
```

In the portal, the app must have user authentication turned on, with
read-and-write permissions and `http://localhost:1410` as a callback URL.

The first time you call one of these functions a browser window opens and
asks you to approve the app. Approve it, come back to R, and the call
carries on. The token is saved to disk under `httr2::oauth_cache_path()`
in a folder named `xapir`, so you sign in once, not every session. When
the token expires it is refreshed in the background; if the refresh fails,
the browser opens again.

To sign in as a different account, delete that folder and call a user
function again:

```{r sign-out}
unlink(file.path(httr2::oauth_cache_path(), "xapir"), recursive = TRUE)
```

## What a call costs

X bills a read per item returned and a write per request, at the prices
on its [pricing page](https://docs.x.com/x-api/getting-started/pricing)
(24 September 2026). A post is US$0.005, a user US$0.010, a follower or
followed account US$0.010, a list, space or community US$0.005, and a
user who liked, muted or blocked something US$0.001. Fields and
expansions are free, which is why every reader asks for every field by
default.

Counts are not free. A count request is billed once, US$0.005 for the
last seven days and US$0.010 for the full archive, however many posts it
covers, so it is still the cheap way to size a query before you pay for
the posts. Trends are US$0.010 a request.

Every reader prints the most it can spend before its first request, using
the `max_posts` (or `max_users`) cap:

```
#> Reading up to 500 posts, about $2.50. Set max_posts to change this.
```

A reader that pages through results also prints what it actually read at
the end, because a pull often stops before the cap:

```
#> Finished getting posts on page 1
#> Finished getting posts on page 2
#> Read 143 posts, about $0.72.
```

A reader billed per request, and every write, prints its price instead:

```
#> This request costs about $0.005.
```

Two rules keep a mistake small. The cap must be a finite number, so
`max_posts = Inf` stops before the first request. And the last page is
trimmed, so a pull never holds more posts than you asked for.

Three details are worth knowing:

* A `username` costs one user read ($0.010) to turn the handle into an id.
  `get_timeline()`, `get_followers()`, `get_owned_list()` and the other
  readers that take a handle also take `user_id`, which skips that read.
  Keep ids as text, `"2244994945"`, because as numbers they lose digits.
* The pricing page de-duplicates reads within a UTC day: reading the same
  post twice on the same day is billed once. Pay-per-use is also capped
  at 3 million post reads a month; `get_usage()` shows where you stand.
* Every price lives in one table. If X changes one, set
  `options(xapir.prices = list(posts = 0.006))` in your `.Rprofile` and
  every message follows. The older `xapir.price_per_post` and
  `xapir.price_per_user` options still work.

## Read, save, unfold

### Read

`get_timeline()` reads an account's posts, newest first. Cap the pull, and
give a window with `start_time` and `end_time` when you want one.
`iso_8601()` turns a date into the format the API wants.

```{r read}
timeline <- get_timeline(
  username   = "Tesla",
  max_posts  = 500,
  start_time = iso_8601("2026-01-01"),
  end_time   = iso_8601("2026-02-01")
)
```

The result is a list of pages, each holding `data`, `includes` and `meta`
as the API returned them. It is not a table yet, on purpose: it holds
everything X sent, including the posts that were quoted or reposted and
the users, media, polls and places they refer to, so you never pay for the
same posts twice.

### Save

Save the raw pull straight away, before you touch it.

```{r save}
saveRDS(timeline, "tesla-2026-01.rds")
```

Every function in the rest of this guide reads from that file. Reading it
back is free; reading the timeline again is not.

```{r load}
timeline <- readRDS("tesla-2026-01.rds")
```

### Unfold

`extract_post()` turns the pages into one row per post with the same 24
columns every time.

```{r extract-post}
post <- extract_post(timeline)
post
```

The columns, in order: `created_at` (UTC; pass `tz = Sys.timezone()` for
local time), `text` (the whole post, even past 280 characters, with
`is_long_post` marking the long ones), `lang`, `possibly_sensitive`,
`paid_partnership` (`TRUE` when the author disclosed the post as paid
promotion), `article_title`, `post_type` (Thread, Post, Quote post, Reply
or Repost), the six counts (`impression_count`, `like_count`,
`repost_count`, `quote_count`, `reply_count`, `bookmark_count`),
`reply_settings`, the ids of the post this one reposted, quoted or replied
to (`reposted`, `quoted`, `replied_to`), `in_reply_to_user_id`, `user_id`,
`community_id` (`NA` unless the post was made in an X community),
`conversation_id`, `post_url` and `post_id`.

On a repost, `like_count`, `reply_count`, `quote_count`, `bookmark_count`
and `repost_count` are `NA`, because those numbers belong to the original
post. Its `impression_count` is its own.

Media come out with `extract_post_media()`: one row per media item per
post, so a post with three photos gives three rows. For a video or an
animated GIF, `url` is the mp4 with the highest bit rate the API offers.

```{r extract-media}
post_media <- extract_post_media(timeline)
post_media
```

The two tables join on `post_id`. Which videos drew the most views?

```{r join}
library(dplyr)

post |>
  select(post_id, created_at, text, like_count) |>
  inner_join(post_media, by = "post_id") |>
  filter(type == "video") |>
  arrange(desc(view_count))
```

## The twelve tables

Each `extract_*()` function reads the saved pull and returns one table.
Together they are the course's data model. None of them calls the API.

| Function                            | One row per                          | Key columns                                         |
|-------------------------------------|--------------------------------------|-----------------------------------------------------|
| `extract_post()`                    | post                                 | `post_id`, `user_id`, `text`, the counts            |
| `extract_user()`                    | user                                 | `user_id`, `username`, `followers_count`            |
| `extract_post_media()`              | media item per post                  | `post_id`, `media_id`, `type`, `url`, `view_count`  |
| `extract_post_url()`                | link per post                        | `post_id`, `expanded_url`, `title`, `image_url`     |
| `extract_post_mention()`            | mentioned account per post           | `post_id`, `username`, `user_id`                    |
| `extract_post_hashtag()`            | hashtag per post                     | `post_id`, `hashtag`                                |
| `extract_post_cashtag()`            | cashtag per post                     | `post_id`, `tag`                                    |
| `extract_post_context()`            | context annotation per post          | `post_id`, `domain_name`, `entity_name`             |
| `extract_post_entity_annotation()`  | named entity per post                | `post_id`, `normalized_text`, `type`, `probability` |
| `extract_post_poll_option()`        | poll option per post                 | `post_id`, `poll_id`, `label`, `votes`              |
| `extract_post_place()`              | tagged post                          | `post_id`, `place_id`, `full_name`, `country`       |
| `extract_post_edited_post_id()`     | earlier version of an edited post    | `post_id`, `edited_post_id`                         |

Eleven of the tables carry `post_id`, so any of them joins to the post
table on that column. `extract_user()` carries `user_id` instead and joins
to the post table's `user_id`, which is the author.

```{r join-user}
post |>
  left_join(extract_user(timeline), by = "user_id") |>
  select(username, created_at, text, like_count)
```

Two properties hold for every table. First, the columns are always the
same, even when there are no rows: a timeline with no polls gives a
zero-row poll table with the usual columns, so a script written on one
pull runs on the next. Second, each table includes the posts a timeline
quotes, replies to or reposts by default, so a quoted post's photo is in
the media table too. Pass `include_referenced_posts = FALSE` to keep only
the posts the endpoint returned. `extract_user()` has no such switch; it
lists every user in the pull.

The other post readers return the same list of pages and feed the same
tables: `get_recent_post()` (a search over the last seven days),
`get_all_post()` (the full archive, on pay-per-use or Enterprise access),
`get_post()` (up to 100 posts by id), `get_quote_post()`, `get_repost()`,
`get_mentions()`, `get_list_posts()`, `get_space_posts()`,
`get_bookmark()`, `get_liked_posts()`, `get_account_timeline()` and
`get_repost_of_me()`. The user readers, such as
`get_users_by_usernames()`, `get_followers()` and `get_list_member()`,
return a tibble directly with the same 24 user columns as
`extract_user()`.

## Followers and lists

Who follows an account? `get_followers()` returns one row per follower
with the 24 user columns, so there is nothing to unfold. It is billed per
user ($0.010), so cap it; the default `max_users = 1000` is about $10.
Pass `user_id` when you know it and the handle lookup is skipped.

```{r followers}
followers <- get_followers(username = "Tesla", max_users = 200)
#> Reading up to 200 users, about $2.00. Set max_users to change this.

followers |>
  arrange(desc(followers_count)) |>
  select(username, name, followers_count, verified)
```

`get_following()` is the same in the other direction, and
`get_reposted_by()` lists who reposted one post.

A list is a curated set of accounts, and its timeline is a post reader
like any other: `get_list_posts()` returns pages, so save them and unfold
them with the same twelve tables. `get_list_by_id()`, `get_list_member()`
and `get_list_followers()` describe the list itself.

```{r list-posts}
list_posts <- get_list_posts(list_id = "1146654567674912769", max_posts = 300)
saveRDS(list_posts, "ev-makers-list.rds")

post <- extract_post(list_posts)
post |>
  left_join(extract_user(list_posts), by = "user_id") |>
  count(username, sort = TRUE)
```

## Writing to X

Every write acts as the account that signed in, so each of these opens
the browser the first time. Each returns the API's `data` invisibly and
stops with X's own message when the request is refused. A rate limit or a
server error is retried, waiting as long as X asks.

```{r write}
# Post, and keep the id so you can delete it later
new_post <- create_post(text = "Hello from R!")
new_post$id

# Delete one or more of your own posts; one row per id
delete_post(post_ids = new_post$id)

# Like, repost, and undo either, by post id
like_post(post_id = "1234567890123456789")
unlike_post(post_id = "1234567890123456789")
create_repost(post_id = "1234567890123456789")
delete_repost(post_id = "1234567890123456789")

# Bookmarks are always your own
create_bookmark(post_id = "1234567890123456789")
delete_bookmark(post_id = "1234567890123456789")

# Follow, unfollow, mute and unmute; the source is the account that signed in
follow_user(source_username = "your_handle", target_username = "XDevelopers")
unfollow_user(source_username = "your_handle", target_username = "XDevelopers")
mute_user(source_username = "your_handle", target_username = "XDevelopers")
unmute_user(source_username = "your_handle", target_username = "XDevelopers")

# Block and unblock; the target can be a handle or an id
block_user(target_username = "spammer")
unblock_user(target_username = "spammer")

# Hide a reply to one of your posts, and show it again
hide_reply(reply_id = "1234567890123456789")
hide_reply(reply_id = "1234567890123456789", hidden = FALSE)

# Lists: create one, add a member, pin it, delete it
new_list <- create_list(name = "EV makers")
add_list_member(list_id = new_list$id, username = "Tesla")
pin_list(list_id = new_list$id)
delete_list(list_id = new_list$id)
```

Every write prints its price before the request. A post is $0.015, but a
post whose text carries a URL is $0.200, more than thirteen times as
much; the price line shows which rate applies, so read it before posting
a batch with links.

To post a photo, GIF or video, upload it first. `upload_media()` sends
the file in chunks, waits for X to finish processing a video, and returns
a media id. Pass that id to `create_post()` as `media_ids`; a post can
carry up to four photos, or one GIF, or one video, and media cannot be
combined with a poll.

```{r media}
media_id <- upload_media("chart.png", alt_text = "Sales by month, 2026")
create_post(text = "Our year so far", media_ids = media_id)

# A reply, a quote and a poll take plain arguments too
create_post(text = "Agreed!", reply_to_post_id = new_post$id)
create_post(text = "Which one?", poll_options = c("This", "That"),
            poll_duration_minutes = 60)
```

`delete_post()` deletes in batches of five with a fifteen-minute pause
between them, which stays inside X's deletion cap on every tier. Raise
`batch_size` on a paid tier to go faster.

## Where next

The [reference index](https://Ivey-Business-School.github.io/xapir/reference/index.html)
lists every function with its arguments and columns, grouped the same way
as this guide. The [release notes](https://Ivey-Business-School.github.io/xapir/news/index.html)
say what changed in each version and which column or argument names
moved.
